---
title: "Tempest calibrations 2019: weight analysis - assumptions"
author: "Liara T'Soni"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
# set global chunk options
knitr::opts_chunk$set(echo = FALSE) # set TRUE if you want to have code in the report
options(knitr.table.format = "html") # set 'latex' if you want to print to pdf
```

```{r libraries, include = FALSE}
library('knitr')
library('tidyverse')
library('kableExtra') # for nice tables
# install.packages("kableExtra")
library('gridExtra') #for plot arrangement
library('DescTools') #for AUC and assumptions calculation
# install.packages('DescTools')
```
  
## Overview  
  
  This report covers all data manipulation performed during data cleaning process and evaluation of variables for assumptions of normal distribution and homogeneity of variance. Data evaluated here are the weight measures. For detailed description of methods used please refer to *project_overview.Rmd*.  
  
```{r data, include = FALSE}
#data <- read_csv('example_project/weight/cleandata/project-weight-cleandata.csv') #for work with chunks
data <- read_csv('cleandata/project-weight-cleandata.csv') #for knitting

#separate data into groups
group_A <- data %>% filter(treatment == 'A')
group_B <- data %>% filter(treatment == 'B')

```

```{r helper_funcs, include = FALSE}
#HELPER FUNCTIONS-------------------------------------------------------------------------------
normal_check <- function(select_data) {
  #checks for normality of the sample, returns W and p-value
  test <- shapiro.test(select_data)
  return(c(W = round(test$statistic, 3),
           p_value = round(test$p.value, 3)))
}

variance_check <- function(dataframe, select_data) {
  #null hypothesis is that all populations variances are equal; the alternative hypothesis is that at least two of them differ
  #checks for homogeneity of variance of the sample, returns df, F value, and p-value
  test <- LeveneTest(select_data ~ treatment, data = dataframe)
  return(c(Df = test$Df,
           F_value = round(test$`F value`[1], 3),
           p_value = round(test$P[1], 3)))
}


data_summary_SE <- function(x) {
  #calculates mean, SE, and interval +- SE
  m <- mean(x)
  se <- sd(x)/sqrt(length(x))
  ymin <- m - se
  ymax <- m + se

  return(c(mean=m, sem=sem, mean_minus_se=ymin ,mean_plus_se=ymax))
}

plot_hist_dens <- function(variable, x_text) {
  #plots histograms and density plots for 2 treatments
  ggplot(data, aes(x = variable, colour = treatment)) + 
  geom_histogram(aes(y = ..density..),
                 fill = 'white',
                 position = 'dodge',
                 bins = 10) +
  geom_density(size = 1) +
  scale_color_manual(values = c('orange', 'blue'),
                     breaks = c('A', 'B')) +
  theme_bw() +
  theme(legend.title = element_blank(), legend.position = "none",
        plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(face = "italic", hjust = 0.5), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        strip.background = element_blank()) +
  labs(x = x_text, 
       y = 'Density')
}


plot_qq <- function(variable, y_text){
  #plots qq plot for 2 treatments
  ggplot(data, aes(sample = variable, colour = treatment)) + 
  geom_point(stat = 'qq') +
  geom_abline(intercept = mean(variable), slope = sd(variable), size = 1) +
  scale_color_manual(values = c('orange', 'blue'),
                     breaks = c('A', 'B')) +
  theme_bw() +
  theme(legend.title = element_blank(), legend.position = "none",
        plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(face = "italic", hjust = 0.5), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        strip.background = element_blank()) +
  labs(x = 'normal', 
       y = y_text)
}

#---
```

  
## Data cleanup  
  
#### First measurement  
  
  Describe conditions based on which data points were included or excluded. 

## Sample size     
  
  Sample size for data per group are as follows:  
```{r sample_size}
sample_size <- data %>% 
  group_by(treatment) %>% 
  summarise(`sample size` = n())

kable(sample_size, escape = FALSE) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)

```
  
## Assumptions check  
  
  Check for normality and homogeneity of variance was performed using Shapiro-Wilk and Levene's test.  *Shapiro-Wilk test* for normality showed that ...(describe result here)... normally distributed.   On the other hand, *Levene's test* for homogeneity of variance showed that ...(describe result here. e.g. all variables have equal variances)... .  Results from assumption tests are summarised below.  
  
  
```{r assumptions_check, warning = FALSE}
#Shapiro test
group_A_norm <- as.data.frame(data.frame(lapply(group_A[2:10], normal_check)) %>% t()) %>% 
  rename(W = W.W) %>% 
  mutate(variable = row.names(.),
         W = W,
         p_value = cell_spec(p_value, bold = ifelse(p_value < 0.05, TRUE, FALSE))) %>% 
  select(variable, W, p_value)

group_B_norm <- as.data.frame(data.frame(lapply(group_B[2:10], normal_check)) %>% t()) %>% 
  rename(W = W.W) %>% 
  mutate(variable = row.names(.),
         W = W,
         p_value = cell_spec(p_value, bold = ifelse(p_value < 0.05, TRUE, FALSE))) %>% 
  select(variable, W, p_value)

data_normal <- left_join(group_A_norm, group_B_norm, by = 'variable')


#display data in table
s_w_table_names <- c('variable','W', 'p value', 'W', 'p value')
kable(data_normal, escape = FALSE, col.names = s_w_table_names) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  add_header_above(c(' ' = 1, 'Group A' = 2, 'Group B' = 2)) %>%  #check if this holds when you change order in df
  add_header_above(c(' ' = 1, 'Shapiro-Wilk test of normality' = 4)) %>% 
  footnote(general_title = 'Bold',
           general = 'p value < 0.05, distribution is not normal',
           footnote_as_chunk = TRUE)

# -------------------------------------------
#Levene test
data_variance <- as.data.frame(data.frame(lapply(data[2:10], variance_check, dataframe = data)) %>% t()) %>% 
  mutate(variable = row.names(.),
         Df1 = Df1,
         Df2 = Df2,
         `F value` = F_value,
         `p value` = cell_spec(p_value, bold = ifelse(p_value < 0.05, TRUE, FALSE))) %>%
  select(variable, Df1, Df2, `F value`, `p value`) %>% 
  unite(Df, Df1, Df2, sep = ', ')
#View(data_variance)

#display data in table
kable(data_variance, escape = FALSE) %>% 
  kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  add_header_above(c(' ' = 1, 'Levene test of homogeneity of variance' = 3)) %>% 
  footnote(general_title = 'Bold',
           general = 'p value < 0.05, homogeneity of variance violated',
           footnote_as_chunk = TRUE)

```
  
  ***  
  
  Below is a set of plots showing distribution of variables of interest and their comparison with normal distribution in qqplot. Blue = Group A, yellow = Group B   
  
```{r histograms}
#create plots for all variables
#group pairs of histograms and qq plots

hist <- plot_hist_dens(data$`2019-03-04`, 'Weight [g]')
qq <- plot_qq(data$`2019-03-04`, 'Weight [g]')
grid.arrange(hist, qq, ncol = 2, top = 'Weight of parts on 2019-03-04')

```
  
  ***  
  
  
  